Mixing time estimation in reversible Markov chains from a single sample path

### 2015-06-02 abstract.tex
minor

%!TEX root =  matrix-est.tex
This article provides the first efficient, non-heuristic procedure for computing a fully data dependent interval
that traps the mixing time of a finite reversible ergodic Markov chain with a prescribed probability.
The interval is computed from a single finite-length sample
path from the Markov chain, and does not require the 
knowledge of any parameters of the chain.
%bounds on the mixing time itself.
This stands in contrast to previous approaches, which either 
only provide point estimates, or require a
reset mechanism, or additional prior knowledge.
%Most notably, 
%the procedure produces a fully empirical confidence
%interval (i.e., it depends only on the sample path and a given
%confidence parameter $\delta$) that is always valid with probability
%$1-\delta$; and 
The width of the interval converges to zero at a
$\sqrt{n}$ rate, where $n$ is the length of the sample path.
Upper and lower bounds are given on the number of samples required
to achieve a constant accuracy estimate in terms of the chain's parameters,
in particular, based on the absolute spectral gap $\gap$.
\todoc{We may need to add some explanation to connect $\gap$ estimates
to $\tmix$ estimates.}
Our lower bounds indicate that, 
unless further restrictions are placed on the chain,
no procedure can achieve a constant
accuracy before seeing each state at least  $\Omega(1/\gap)$ times.
Future directions of research are identified.
\todoc{Such as: closing the gap. Extension to large state spaces, 
based on additional assumptions. Even if $\gap$ is large, we need many samples.
Maybe we need even stronger assumptions, like the decay rate of eigenvalues.
}
%If $\gap$ denotes the absolute spectral gap of the unknown Markov chain
%and the chain's stationary distribution is minorized by $\pimin$,
%our procedure requires $\tilde{O}( \frac{1}{\pimin\gap^3})$ observations
% to estimate the mixing time up to a constant multiplicative accuracy, 
%while we also establish that any algorithm to achieve the same
%goal would need $\Omega( d\log(d)/\gap + 1/\pimin)$ samples.

>>>>

%!TEX root =  matrix-est.tex
This article provides the first efficient, non-heuristic procedure for computing a fully data dependent interval
that traps the mixing time of a finite reversible ergodic Markov chain with a prescribed probability.
The interval is computed from a single finite-length sample
path from the Markov chain, and does not require the 
knowledge of any parameters of the chain.
%bounds on the mixing time itself.
This stands in contrast to previous approaches, which either 
only provide point estimates, or require a
reset mechanism, or additional prior knowledge.
%Most notably, 
%the procedure produces a fully empirical confidence
%interval (i.e., it depends only on the sample path and a given
%confidence parameter $\delta$) that is always valid with probability
%$1-\delta$; and 
The width of the interval converges to zero at a
$\sqrt{n}$ rate, where $n$ is the length of the sample path.
Upper and lower bounds are given on the number of samples required
to achieve a constant accuracy estimate in terms of the chain's parameters,
in particular, based on the absolute spectral gap $\gap$.
\todoc{We may need to add some explanation to connect $\gap$ estimates
to $\tmix$ estimates.}
The lower bounds indicate that, 
unless further restrictions are placed on the chain,
no procedure can achieve a constant
accuracy before seeing each state at least  $\Omega(1/\gap)$ times.
Future directions of research are identified.
\todoc{Such as: closing the gap. Extension to large state spaces, 
based on additional assumptions. Even if $\gap$ is large, we need many samples.
Maybe we need even stronger assumptions, like the decay rate of eigenvalues.
}
%If $\gap$ denotes the absolute spectral gap of the unknown Markov chain
%and the chain's stationary distribution is minorized by $\pimin$,
%our procedure requires $\tilde{O}( \frac{1}{\pimin\gap^3})$ observations
% to estimate the mixing time up to a constant multiplicative accuracy, 
%while we also establish that any algorithm to achieve the same
%goal would need $\Omega( d\log(d)/\gap + 1/\pimin)$ samples.

### 2015-06-02 rates.tex
minor

Our upper bound is demonstrated using estimators based on natural
plug-in estimates.
We also provide an estimator for the minimum stationary probability
in the process, so that one can use the earlier stated bounds to trap
the mixing time.

>>>>

Our upper bound is demonstrated using estimators based on natural
plug-in estimates.
We also provide an estimator for the minimum stationary probability
in the process, so that one can use the bounds from
\cref{eq:mixing-time-bound} to trap
the mixing time.
